Bootstrap learning via modular concept discovery

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
Edits to abstract and introduction.

\maketitle
\begin{abstract}
%\small
 Suppose a learner is faced with a domain of problems about which it
 knows nearly nothing. It does not know the distribution of problems,
 the space of solutions is not smooth, and the reward signal is
 uninformative, providing perhaps a few bits of information but not
 enough to steer it effectively. How can such a learner ever get off
 the ground? A common intuition is that if the solutions to these
 problems share a common structure, and the learner can solve some
 simple problems by brute force, it should be able to extract useful
 components from these solutions and, by composing them, explore the
 solution space more efficiently. Here, we formalize this intuition,
 where the solution space is the space of typed functional programs
 and the gained information is stored as a stochastic grammar over
 programs. We propose an iterative procedure for exploring such
 spaces: in the first step of each iteration, the learner explores a
 finite subset of the domain, guided by a stochastic grammar; in the
 second step, the learner compresses the successful solutions from the
 first step to estimate a new stochastic grammar. We test this
 procedure on symbolic regression and boolean circuit learning and
 show that the learner discovers modular concepts for these
 domains. Whereas the learner is able to solve almost none of the
 posed problems in the procedure's first iteration, it rapidly becomes
 able to solve a large number by gaining abstract knowledge of the
 structure of the solution space.

@@

\section{Introduction}
A central question for AI is how to create intelligent agents without
building in vast amounts of pre-specified knowledge that is tailored
to one domain or another. Any attempt to solve this problem must
contend with the fact that the less an agent knows about solving
problems in a domain, the worse its attempts are, and the less it
learns about how to solve subsequent problems. For example, imagine
you know nothing about baking but are supplied with a stocked kitchen:
your initial attempts will be uniformly disastrous, and you will learn
little from them, except that you did not create a baked good. A
little bit of knowledge about baking would go a long way towards
making your initial attempts more informative.

Like many of the problems in AI, this one can be cast as a search
problem. There are many search algorithms, but they typically exploit
local or topological structure in the search space. For example,
convex optimization and gradient based search methods take
advantage of local smoothness to move in good directions. Local search
algorithms move from one candidate solution to a neighboring
one. Heuristic search techniques take advantage of topological
structure to prune off large parts of the solution space.

The absense of domain specific knowledge makes exploiting structure
difficult, especially if the structure of the search space is not
smooth. This is particularly true of problems -- such as discovering
scientific theories, developing new technologies, learning how to bake
-- whose solutions can be naturally framed as programs. Here we ask
how to bootstrap domain specific knowledge and extract structure from
the space of programs by attempting to solve multiple related problems
in a given domain at once.

Out approach is inspired by the programmer who -- having reused
several subroutines to solve different but similar simple tasks --
creates a function library to help manage the complexity of larger
projects. The E.C. algorithm, which we introduce here, builds a library
of reusable program components and places a distribution over these,
effectively learning from simple tasks a search space structure that
enables it to solve more complex tasks.

@@

\section{Related Work}
The idea of discovering and using reusable subcomponents is part of an
old tradition in AI. For instance, it is one of the central themes in
Herb Simon's 1969 ``The Sciences of the
Artificial''~\cite{simon1996sciences}. Rendell's 1985
paper~\cite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
Induction Using Layered Information Compression: Tractable Feature
Formation in Search,'' presented within a classical search paradigm
the idea that compression provides a heuristic for constructing good
representations. Within the Inductive Logic Programming (ILP)
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
was also applied in the multitask
setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
learning functional programs, and believe that the modularity and
compositionality afforded by this representation is necessary for
these ideas to be successful.

@@

\begin{figure}
%%\vspace{-.3cm}
\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
%%\vspace{-.6cm}
\caption{Distribution of boolean functions among (a) the set of 1000
  circuits randomly constructed from AND, OR and NOT gates (see text);
  b) the first 1000 expressions from the grammar over expressions
  \emph{before} learning; c) the first 1000 expressions from the
  grammar over expressions \emph{after} learning. Gray levels indicate
  percent of instances (circuits/expressions) which evaluate to that
  function.\label{fig:booldistr}}
%%\vspace{-.4cm}
\end{figure}

>>>>

\maketitle
\begin{abstract}
%\small
 Suppose a learner is faced with a domain of problems about which it
 knows nearly nothing. It does not know the distribution of problems,
 the space of solutions is not smooth, and the reward signal is
 uninformative, providing perhaps a few bits of information but not
 enough to steer the learner effectively. How can such a learner ever
 get off the ground? A common intuition is that if the solutions to
 these problems share a common structure, and the learner can solve
 some simple problems by brute force, it should be able to extract
 useful components from these solutions and, by composing them,
 explore the solution space more efficiently. Here, we formalize this
 intuition, where the solution space is the space of typed functional
 programs and the gained information is stored as a stochastic grammar
 over programs. We propose an iterative procedure for exploring such
 spaces: in the first step of each iteration, the learner explores a
 finite subset of the domain, guided by a stochastic grammar; in the
 second step, the learner compresses the successful solutions from the
 first step to estimate a new stochastic grammar. We test this
 procedure on symbolic regression and boolean circuit learning and
 show that the learner discovers modular concepts for these
 domains. Whereas the learner is able to solve almost none of the
 posed problems in the procedure's first iteration, it rapidly becomes
 able to solve a large number by gaining abstract knowledge of the
 structure of the solution space.

@@

\section{Introduction}
A central question for AI is how to create intelligent agents without
building in vast amounts of pre-specified knowledge that is tailored
to one domain or another. Any attempt to solve this problem must
contend with the fact that the less an agent knows about solving
problems in a domain, the worse its attempts are, and, therefore, the
less informative they are about how to solve subsequent problems. For
example, imagine you know nothing about baking but are supplied with a
stocked kitchen: your initial attempts will be uniformly inedible, and
you will learn little from them. If, on the other hand, you know the
basics, you are far more likely to generate recipes that are
somewhat successful and on whose basis you might improve your skills.

Like many of the problems in AI, this one can be cast as a search
problem: the goal is to find a baking procedure that maximizes certain
criter.  There are many search algorithms, but they typically exploit
local or topological structure in the search space. For example, local
search algorithms move from one candidate solution to a neighboring
one. Heuristic search techniques take advantage of topological
structure to prune off large parts of the solution space.

Domain specific knowledge provides this structure. In local search, it
defines the similarity metric between solutions; in heuristic search
it provides the heuristic.

Here we ask how to bootstrap domain specific knowledge and extract
structure from the space of possible solutions by attempting to solve
multiple related problems in a given domain at once. Our approach is
motivated by the observation that many interesting problems -- such as
discovering scientific theories, developing new technologies, learning
to play a new instrument -- have solutions that are natrually
specified as programs.

Our proposed solution is thus inspired by the programmer who -- having
reused subroutines to solve similar simple tasks -- creates a function
library to help manage the complexity of larger projects. The
E.C. algorithm, which we introduce here, builds a library of reusable
program components and places a distribution over these, effectively
learning from simple tasks a search space structure that enables it to
solve more complex tasks.

@@

\section{Related Work}
The idea of discovering and using reusable subcomponents is part of an
old tradition in AI. For instance, it is one of the central themes in
Herb Simon's 1969 ``The Sciences of the
Artificial''~\cite{simon1996sciences}. Rendell's 1985
paper~\cite{DBLP:conf/ijcai/Rendell85}, ``Substantial Constructive
Induction Using Layered Information Compression: Tractable Feature
Formation in Search,'' presented within a classical search paradigm
the idea that compression provides a heuristic for constructing good
representations. Within the Inductive Logic Programming (ILP)
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. Later, predicate invention
was also applied in the multitask
setting~\cite{DBLP:conf/ilp/KhanMP98}. We apply these ideas to
learning functional programs and believe that the modularity and
compositionality afforded by this representation is necessary for
these ideas to be successful.

@@

\begin{figure}
%%\vspace{-.3cm}
\includegraphics[width=\linewidth]{figures/circuitDistrComp2.png}
%%\vspace{-.6cm}
\caption{Distribution of boolean functions among (a) the set of 1000
  circuits randomly constructed from AND, OR and NOT gates (see text);
  b) the first 1000 expressions from the grammar over expressions
  \emph{before} learning; c) the first 1000 expressions from the
  grammar over expressions \emph{after} learning. Gray levels indicate
  percent of instances (circuits/expressions) which evaluate to that
  function.\label{fig:booldistr}}
%%\vspace{-.4cm}
\end{figure}
